\chapter*{Acknowledgements}
First and foremost, my gratitude goes out to Alexandre Gramfort: my advisor, mentor and friend. Through these years, Alex has inspired me with his technical knowledge, his vision for long term impact, philosophy of open science, and wisdom. He introduced me to the welcoming and progressive MNE community which ultimately formed my collaboration network. It is these interactions that often led to new research projects and ideas. A prime example of this is my work on \emph{autoreject} which developed out of my collaboration with Denis Engemann, a core contributor to MNE. It has been a pleasure working with Denis, as he shared his knowledge on the subtleties of MEG signal analysis and motivated me through our mini coding sprints. I would like to express my heartfelt thanks to Riitta Hari, Lauri Parkkonen, and Pavan Ramkumar who introduced me to the world of machine learning in neuroscience, and offered words of support and encouragement over the years. 

My gratitude goes out to Matti Hamalainen and Stephanie Jones for inviting me to Boston on our collaborative project. I thank my friend and former colleague, Teon Brooks for working with me on the BIDS project and on the realtime module in MNE, Eric Larson for sharing his extensive experience in MEG and signal processing, and the rest of the MNE team for making me believe in the power of open source. I am also grateful to Chris Gorgolewski and Russ Poldrack for inviting me to the BIDS coding sprint at Stanford. The visit was a very stimulating experience for me as well as a unique networking opportunity. I thank my advisor Alex, and Robert Gower for allowing me to take their optimization course, and for being incredibly patient with my questions. I am sure the skills I learned here will be useful for many years to come.

To my co-authors, I thank them for being so tolerant and supportive in our common projects. None of my papers could have been published without the teamwork and efforts put in by them. I am grateful to Umut \c{S}im\c{s}ekli and Tom Dupr\'{e} la Tour for spending their weekends and evenings with me to push forward our sparse coding project. I have benefited a lot from my collaboration with Jaakko Leppakangas who is perhaps one of the most efficient and productive engineers I have met. I am particularly indebted to my colleague, Yousra Bekhti for discussing tricky bugs at work, but more importantly for easing my difficulties with the French administration and with translation. I am grateful to Jean-Baptiste Schiratti for many stimulating technical conversations, and also for helping me with constructive feedback on my writing.

I thank my office mates for sharing the weekly seminars, the Monday cakes, and the latest free food event. I thank my collaborators and friends at the Neurospin laboratory for inviting me to their social events. It is here that I made many like-minded international friends. To my French teacher Françoise and our Tuesday French lunch group, I am immensely indebted. I thank my international friends: Bianca, David, Fosca, Gabriela, Magdalena, Sokhany, and Sophie. I thank my neighbours, my climbing partners, hiking group, and the \textit{desi} Indian community for making me feel home in France: Aakanksha, Anshuman, Chirag, Nilesh, Niraj, Pratheeban, Praveer, Raghav, Shabbir, Sidharth, and Vamsi.

All of this work would not have been possible without my family. My sister continues to surprise and inspire me with her sense of humour and adventure. Finally, I thank my parents for their continued moral support and motivation.

\vspace{3em}
April 12, 2018 \\
Paris
